---
title: "Is this a good model quiz?"
author: "Adi Sarid / adi@sarid-ins.co.il"
date: "July 2019"
output: html_document
---

## Neither good nor bad situation

We're trying to predict $\sin(x)$ using just $x$, i.e.:
The "true" function is 
\[
y = sin(x) + \epsilon
\]

In this case, $\epsilon\sim N(\mu = 0, \sigma = 0.15)$.

```{r}
suppressMessages(suppressWarnings(library(tidyverse)))
set.seed(0)
data_in_partial <- tibble(x = seq(0, pi/2, length.out = 180)) %>% 
  mutate(true_y = sin(x),
         measured_y = sin(x) + rnorm(length(x), 0, 0.15))

lm_y_partial <- lm(data = data_in_partial, formula = measured_y ~ x)

summary(lm_y_partial)
```

This is a great opportunity to talk about the `summary` function when used on a linear regression model.

We have a number of elements to discuss:

   * Residuals
   * p-values (significance testing)
   * R-squared
   
# The nominal prediction

Note how we're also extrapolating for the range $(\frac{\pi}{2}, \frac{3\pi}{2})$. That's what's wrong with the chart.

```{r}
# This is the nominal prediction, while hiding my "true" intentions

data_in_partial %>% 
  mutate(lm_y = predict(lm_y_partial)) %>% 
  select(x, measured_y, lm_y) %>% 
  bind_rows(tibble(x = seq(pi/2, pi/1.5, length.out = 90)) %>% 
              mutate(lm_y = predict(lm_y_partial, .))) %>% 
  gather(type, y, -x) %>% 
  ggplot(aes(x = x, y = y, color = type)) + 
  geom_point() + 
  theme_bw()

```

## How wrong can this go on?

Wrong. Very Wrong.

```{r}

data_in <- tibble(x = seq(0, 2*pi, length.out = 360)) %>% 
  mutate(true_y = sin(x),
         measured_y = sin(x) + rnorm(length(x), 0, 0.15)) %>% 
  mutate(lm_y = predict(lm_y_partial, newdata = .))

data_in %>% 
  gather(type, y, -x) %>% 
  ggplot(aes(x = x, y = y, color = type)) + 
  geom_point() + 
  theme_bw()

```

## Fixing the issue

A linear model, but with a twist.

```{r}

lm_y_sine <- lm(data = data_in_partial, formula = measured_y ~ sin(x))

summary(lm_y_sine)

data_in_w_sine <- tibble(x = seq(0, 2*pi, length.out = 360)) %>% 
  mutate(true_y = sin(x),
         measured_y = sin(x) + rnorm(length(x), 0, 0.15)) %>% 
  mutate(lm_y = predict(lm_y_partial, newdata = .),
         lm_y_sin = predict(lm_y_sine, newdata = .))

data_in_w_sine %>% 
  select(-true_y, -measured_y) %>% 
  gather(type, y, -x) %>% 
  ggplot(aes(x = x, y = y, color = type)) + 
  geom_line(size = 1.5) + 
  theme_bw()

```